---
title: "p8105_hw2_hl3464"
author: "Hongtong Lin"
date: "2026-10-05"
output: github_document
---

load the libraries  
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```
  
## Problem 1  

```{r read csv}
nyc_transit <- 
  read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           na = c("NA", "")) |> 
  janitor::clean_names()
```
  
```{r clean data set}
nyc_transit_tidy <-
  nyc_transit |> 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada
  ) |> 
  mutate(
    entry = recode_values(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
``` 
  
  
## Problem 2  

tidy mr_trash_wheel data frame
```{r tidy mr_trash_wheel}
mr_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Mr. Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:N")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    sports_balls = round(sports_balls, 0) |> 
      as.integer(),
    year = as.integer(year),
    trash_wheel = "Mr. Trash Wheel"
  ) |> 
  filter(!is.na(dumpster))
```
  
tidy prof_trash_wheel data frame
```{r tidy prof_trash_wheel}
prof_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Professor Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:M")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  ) |> 
  filter(!is.na(dumpster))
```
  
tidy gwynns_trash_wheel data frame
```{r tidy gwynns_trash_wheel}
gwynns_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Gwynns Falls Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:L")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    year = as.integer(year),
    trash_wheel = "Gwynns Trash Wheel"
  ) |>
  filter(!is.na(dumpster))
```
  
Join data frames together  
```{r}
trash_wheel <- 
  bind_rows(mr_trash_wheel, prof_trash_wheel, gwynns_trash_wheel) |> 
  relocate(trash_wheel)

head(trash_wheel)
```

```{r, eval = FALSE}
str(trash_wheel)  

prof_trash_wheel |> 
  pull(weight_tons) |> 
  sum() # 282.26 tons

gwynns_trash_wheel |> 
  filter(
    year == 2022,
    month == "June"
  ) |> 
  pull(cigarette_butts) |> 
  sum() #18120 

```
- There are 15 variables or columns and 1,188 observations (rows) in the trash_wheel data frame. Some key variables are "trash_wheel" for the name of the trash_wheel, "dumpster" as the dumpster number, "weight_tons", and so on.
- The total weight of trash collected by Prof. Trash Wheel is 282.26 tons.
- The total number of cigarette butts collected by Gwynnda in June of 2022 is 18,120.


