---
title: "p8105_hw2_hl3464"
author: "Hongtong Lin"
date: "2026-10-05"
output: github_document
---

load the libraries  
```{r load libraries, message=FALSE}
library(tidyverse)
```
  
## Problem 1  

```{r read csv}
nyc_transit <- 
  read_csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           na = c("NA", "")) |> 
  janitor::clean_names()
```
  
```{r clean data set}
nyc_transit_tidy <-
  nyc_transit |> 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada
  ) |> 
  mutate(
    entry = recode_values(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
``` 
  
  
## Problem 2  

tidy mr_trash_wheel data frame
```{r tidy mr_trash_wheel}
mr_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Mr. Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:N")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    sports_balls = round(sports_balls, 0) |> 
      as.integer(),
    year = as.integer(year),
    trash_wheel = "Mr. Trash Wheel"
  ) |> 
  filter(!is.na(dumpster))
```
  
tidy prof_trash_wheel data frame
```{r tidy prof_trash_wheel}
prof_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Professor Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:M")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  ) |> 
  filter(!is.na(dumpster))
```
  
tidy gwynns_trash_wheel data frame
```{r tidy gwynns_trash_wheel}
gwynns_trash_wheel <- 
  readxl::read_xlsx("202509 Trash Wheel Collection Data.xlsx",
                    sheet = "Gwynns Falls Trash Wheel",
                    skip = 1,
                    range = readxl::cell_cols("A:L")
                    ) |> 
  janitor::clean_names() |> 
  mutate(
    year = as.integer(year),
    trash_wheel = "Gwynns Trash Wheel"
  ) |>
  filter(!is.na(dumpster))
```
  
Join data frames together  
```{r join data frames together, results = "hide"}
trash_wheel <- 
  bind_rows(mr_trash_wheel, prof_trash_wheel, gwynns_trash_wheel) |> 
  relocate(trash_wheel)

head(trash_wheel)
```

```{r check trash_wheel joined data frame, message = FALSE}
str(trash_wheel)
```


```{r calculations from the trash_wheel df, results = "hide", message = FALSE}
prof_trash_wheel |> 
  pull(weight_tons) |> 
  sum() # 282.26 tons

gwynns_trash_wheel |> 
  filter(
    year == 2022,
    month == "June"
  ) |> 
  pull(cigarette_butts) |> 
  sum() #18120 

```
- There are 15 variables or columns and 1,188 observations (rows) in the trash_wheel data frame. Some key variables are "trash_wheel" for the name of the trash_wheel, "dumpster" as the dumpster number, "weight_tons", and so on.
- The total weight of trash collected by Prof. Trash Wheel is 282.26 tons.
- The total number of cigarette butts collected by Gwynnda in June of 2022 is 18,120.
  
  
## Problem 3  

### MCI_baseline data exploration
```{r tidy the mci_baseline df, message = FALSE}
MCI_baseline <- 
  read_csv("data_mci/MCI_baseline.csv",
           skip = 1,
           na = ".") |> 
  janitor::clean_names() |> 
  filter(!is.na(age_at_onset)) |> 
  mutate(
    sex = recode_values(
      sex,
      0 ~ "female",
      1 ~ "male"
    ),
    apoe4 = recode_values(
      apoe4,
      0 ~ "non-carrier",
      1 ~ "carrier"
    )
  )
```
  
```{r overview of the mci_baseline df, results = "hide", message = FALSE}
dim(MCI_baseline) #97 rows, 6 columns after tidying up the data  

read_csv("data_mci/MCI_baseline.csv",
         skip = 1) |> 
  nrow() # 483 rows in the original data set 

mean(MCI_baseline$age_at_onset) # 70.26 years - baseline onset age
```


```{r proportion of carrier in women}
MCI_baseline |> 
  filter(sex == "female") |> 
  summarise(prop_apoe4_carrier = mean(apoe4 == "carrier")) # prop is 0.652
  
```

- The mci_baseline original data set has a lot of missing values represented by "." meaning that individuals are free of MCI at baseline during the follow up. I excluded these missing data because these individuals did not meet the stated inclusion criteria. The original data set also has a row of explanations at the top, so I skipped this line when importing the csv file. After that, I re-coded the values for the sex and apoe4 variables because using numbers 1 and 0 can be confusing.
- 483 individuals were recruited, and 97 of them develop MCI. 
- The average baseline onset age is 70.26 years. 
- 65.2% of women who developed MCI at baseline are apoe4 carriers.  
  

### mci_amyloid data exploration  

```{r tidy up the mci_amyloid df, warning = FALSE, message= FALSE}
mci_amyloid <-
  read_csv("data_mci/mci_amyloid.csv",
           skip = 1,
           na = "NA") |>
  janitor::clean_names() |> 
  mutate(
    baseline = as.numeric(baseline),
    time_2 = as.numeric(time_2),
    time_4 = as.numeric(time_4),
    time_6 = as.numeric(time_6),
    time_8 = as.numeric(time_8)
  ) |> 
  rename(id = study_id) # diectly replace the variable name "study_id" into "id"
```
  
- This mci_amyloid csv file has a top row with some explanations, so i skipped it when importing the csv. There are many NA values in the data set, so I used na = "NA" to identify those values. After importing, I found that almost all variables were in the character data type, so I coerced them into numeric variables.  

no match between 2 data frames  
```{r use anti_join to find ppl only in MCI_baseline}
only_baseline <- anti_join(MCI_baseline, mci_amyloid, by = "id")
nrow(only_baseline) # 14,49,268
```
  
```{r se anti_join to find ppl only in mci_amyloid}
only_amyloid <- anti_join(mci_amyloid, MCI_baseline, by = "id") 
nrow(only_amyloid)
# 393 individuals only in mci_amyloid
```

- Individual with id 14, 49, and 268 are only in the "MCI_baseline" data set.
- 393 individuals in total are only in "mci_amyloid" data set.  

```{r combine 2 data sets: only participants who appear in both data sets}
mci_combined <- inner_join(MCI_baseline, mci_amyloid, by = "id")

nrow(mci_combined) # 94 individuals

write_csv(mci_combined, "data_mci/mci_combined.csv")
```
- 94 individuals in total are found in both the mci_baseline and mci_amyloid data sets. 

